### 2020-07-07 manuscript/pnas_style/manuscript.tex
removed supp materials links

%\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
\templatetype{pnasresearcharticle} % Choose template

\RequirePackage{xr}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\flt}{\mathrm{flat}}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}

\newcommand{\vb}[1]{{\it \color{red} #1}}

% \addbibresource{biblio.bib}

\externaldocument{supp}

\title{Estimating the genome-wide contribution of selection to temporal allele frequency change}

\author[a,b,c,1]{Vince Buffalo}
\author[b,c]{Graham Coop}
\affil[a]{\footnotesize Population Biology Graduate Group, University of California, Davis, 95616}
\affil[b]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}

\leadauthor{Buffalo}

\authorcontributions{V.B. and G.C. designed research and wrote paper; V.B. performed research, contributed new reagents/analytic tools, and analyzed data.}
\authordeclaration{The authors declare no conflict of interest.}
% \equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: vsbuffalo@gmail.com}

\keywords{Linked Selection $|$ Adaptation $|$ Experimental Evolution }

\begin{abstract}
  Rapid phenotypic adaptation is often observed in natural populations and
  selection experiments. However, detecting the genome-wide impact of this
  selection is difficult, since adaptation often proceeds from standing
  variation and selection on polygenic traits, both of which may leave faint
  genomic signals indistinguishable from a noisy background of genetic drift.
  One promising signal comes from the genome-wide covariance between allele
  frequency changes observable from temporal genomic data, e.g.
  evolve-and-resequence studies. These temporal covariances reflect how
  heritable fitness variation in the population leads changes in allele
  frequencies at one timepoint to be predictive of the changes at later
  timepoints, as alleles are indirectly selected due to remaining associations
  with selected alleles. Since genetic drift does not lead to temporal
  covariance, we can use these covariances to estimate what fraction of the
  variation in allele frequency change through time is driven by linked
  selection. Here, we reanalyze three selection experiments to quantify the
  effects of linked selection over short timescales using covariance among
  time-points and across replicates. We estimate that at least 17\% to 37\% of
  allele frequency change is driven by selection in these experiments. Against
  this background of positive genome-wide temporal covariances we also identify
  signals of negative temporal covariance corresponding to reversals in the
  direction of selection for a reasonable proportion of loci over the time
  course of a selection experiment. Overall, we find that in the three studies
  we analyzed, linked selection has a large impact on short-term allele
  frequency dynamics that is readily distinguishable from genetic drift.
\end{abstract}

\begin{document}
\maketitle

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% \linenumbers

\significancestatement{
A long-standing problem in evolutionary biology is to understand the processes
that shape the genetic composition of populations. In a population without
migration, the two processes that change allele frequencies are selection,
which increases beneficial alleles and removes deleterious ones, and genetic
drift which randomly changes frequencies as some parents contribute more or
less alleles to the next generation. Previous efforts to disentangle these
processes have used genomic samples from a single timepoint and models of how
selection affects neighboring sites (linked selection). Here, we use genomic
data taken through time to quantify the contributions of selection and drift to
genome-wide frequency changes. We show selection acts over short timescales in
three evolve-and-resequence studies and has a sizable genome-wide impact.}

\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
roles of genetic drift and selection in shaping genome-wide allele frequency
changes.  Selection can affect allele frequencies, both directly and
indirectly, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome, e.g. linked selection
\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
This approach has shown linked selection has a major role in shaping patterns
of genome-wide diversity across the genomes of a range of sexual species
\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over short
time intervals and of its full impact on genome-wide allele frequency changes.

There are numerous examples of rapid phenotypic adaptation
\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature of
fitness makes detecting the impact of selection on genome-wide variation over
short timescales in sexual populations remarkably difficult \cite{Latta1998-me,
Pritchard2010-tk,Kemper2014-bx}. This is because the effect of selection on a
polygenic trait (such as fitness) is distributed across numerous loci. This can
lead to subtle allele frequency shifts on standing variation that are difficult
to distinguish from background levels of genetic drift and sampling variance.
Increasingly, genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large-effect selected loci \cite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous
paper, we proposed that one useful signal for understanding the genome-wide
impact of polygenic linked selection detectable from temporal genomic data is
the temporal autocovariance (i.e. covariance between two timepoints) of
allele frequency changes \cite{Buffalo2019-io}. These covariances are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are expected to be zero for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a time interval
into the variances and covariances in allele frequency change between
generations.  Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an artificial
selection experiment \cite{Castro2019-uk}. These sequencing selection
experiments have started to uncover selected loci contributing to the adaptive
response; however it is as yet far from clear how much of genome-wide
allele frequency changes are driven by selection or genetic drift. We
repeatedly find a signal of temporal covariance, consistent with linked
selection acting to significantly perturb genome-wide allele frequency changes
across the genome in a manner that other approaches would not be able
differentiate from genetic drift. We estimate a lower bound of the fraction of
variance in allele frequency change caused by selection, as well as the
correlation between allele frequency changes between replicate populations
caused by convergent selection pressures. Overall, we demonstrate that linked
selection has a powerful role in shaping genome-wide allele frequency changes
over very short timescales in experimental evolution.

\section*{Results}

\begin{table*}
  \centering
  %\footnotesize
  \caption{Summary of the main selection studies we analyzed} 
  \centering
  \begin{tabular}{l c c c c c c}
    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size}$^{\dagger}$ & \textbf{Gens.} & \textbf{Timepoints} \\ 
  \hline
    Kelly and Hughes (2019)  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    Barghi et al. (2019) & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    Castro et al. (2019) & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
  \bottomrule
  \end{tabular}

  \addtabletext{$^{\dagger}$Approximate census population size during experiment.}
  \label{table:study_summary} 
\end{table*}

We first analyzed the dataset of Barghi et al. (2019), an evolve-and-resequence
study with ten replicate populations exposed to a high temperature lab
environment and evolved for 60 generations, and sequenced every ten
generations. Using the seven timepoints and ten replicate populations, we
estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
for each of the ten replicates. Each row of these matrices represent the
temporal covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the
allele frequency change (in ten-generation intervals, denoted $\Delta_{_{10}}
p_t$) of some initial reference generation $s$ (the row of the matrix), and
some later timepoint $t$ (the column of the matrix). We corrected these
matrices for biases created due to sampling noise, and normalized the entries
for heterozygosity (see SI Appendix, Sections S1.2 and S1.4). These covariances
are expected to be zero when only drift is acting, as only heritable variation
for fitness can create covariance between allele frequency changes in a closed
population \cite{Buffalo2019-io}.  Averaging across the ten replicate temporal
covariances matrices, we find temporal covariances that are statistically
significant (95\% block bootstraps CIs do not contain zero), consistent with
linked selection perturbing genome-wide allele frequency changes over very
short time periods.  The covariances between all adjacent time intervals are
positive and then decay towards zero as we look at more distant time intervals
(Figure \ref{fig:figure-1} A), as expected when directional selection affects
linked variants' frequency trajectories until ultimately linkage disequilibrium
and the associated additive genetic variance for fitness decays (which could
occur as a population reaches a new optimum, and directional selection weakens)
\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
this general pattern holds; see SI Appendix, Fig. S23.

Since our covariances are averages over loci, the covariance estimate could be
strongly affected by a few outlier regions. To test whether large outlier
regions drive the genome-wide signal we see in the Barghi et al. (2019) data,
we calculate the covariances in 100kb windows along the genome (we refer to
these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates (SI
Appendix, Table S1 and Fig. S24) confirm the patterns we see using the mean
covariance, establishing that genomic temporal covariances are non-zero due to
the impact of selection acting across many genomic regions.

\begin{figure*}
  \centering
  \includegraphics[width=15cm]{figure-1-pnas.pdf}

  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
    populations, through time from the Barghi et al. (2019) study. Each line
    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
    reference generation $s$ to a later time $t$ which varies along the x-axis;
    each line corresponds to a row of the upper-triangle of the temporal
    covariance matrix with the same color (upper right). The ranges around each
    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: A lower
    bound on the proportion of the total variance in allele frequency change
    explained by linked selection, $G(t)$, as it varies through time $t$ along
    the x-axis. The black line is the $G(t)$ averaged across replicates, with
    the $95\%$ block-bootstrap confidence interval. The other lines are the
    $G(t)$ for each individual replicate, with colors indicating what subset of
    the temporal-covariance matrix to the right is being included in the
    calculation of $G(t)$.}

  \label{fig:figure-1}
\end{figure*}

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change: the
ratio of total covariance in allele frequency change to the total variance in
allele frequency change. We denote the change in allele frequency as $\Delta
p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
Since the total variation in allele frequency change can be partitioned into
variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
\Delta p_j)$ (we correct for biases due to sequencing depth), and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}.
\end{align}
%
This estimates the impact of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. When the sum of the covariances is
positive, this measure can intuitively be understood as a lower bound on
relative fraction of allele frequency change normally thought of as ``drift''
that is actually due to selection. Additionally, $G(t)$ can be understood as a
short-timescale estimate of the reduction in neutral diversity due to linked
selection (or equivalently the reduction in neutral effective population size
needed to account linked selection, see SI Appendix, Section S7). Since the
Barghi et al. (2019) experiment is sequenced every ten generations, the
numerator uses the covariances estimated between ten-generation blocks of
allele frequency change; thus the strong, unobservable, covariances between
adjacent generations do not contribute to the numerator of $G(t)$. Had these
covariances been measurable on shorter timescales, their cumulative effect
would likely have been higher yet (see SI Appendix, Sections S2 and
S8.4 for more details).  Additionally, selection inflates the variance in
allele frequency change per generation; however, this effect cannot be easily
distinguished from drift. For both these reasons, our measure $G(t)$ is quite
conservative (we demonstrate this through simulations in SI Appendix, Section
S8.4).  Still, we find a remarkably strong signal. Greater than $20\%$ of
total, genome-wide allele frequency change over 60 generations is the result of
selection (Figure \ref{fig:figure-1} B). This proportion of variance
attributable to selection builds over time in Figure \ref{fig:figure-1}B as the
effects of linked selection are compounded over the generations unlike genetic
drift. Our G(t) starts to plateau to a constant level as the covariances from
earlier generations have decayed and so no longer contribute as strongly
(Figure \ref{fig:figure-1}).

Additionally, we looked for a signal of temporal autocovariance in Bergland et
al. (2014), a study that collected \emph{Drosophila melanogaster} through
Spring-Fall season pairs across three years. If there was a strong  pattern of
genome-wide fluctuating selection, we might expect a pattern of positive
covariances between similar seasonal changes, e.g.  Spring-Fall in two adjacent
years, and negative covariances between dissimilar seasonal changes, e.g.
Spring-Fall and Fall-Spring in two adjacent years.  However, we find no such
signal over years, and in reproducing their original analysis, we find that
their number of statistically significant seasonal SNPs is not enriched
compared to an empirical null distribution created by permuting seasonal
labels; we discuss this in more depth in SI Appendix, Section S6).

The replicate design of Barghi et al. (2019) allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are expected to
have allele frequency changes in the same direction if the selection pressures
are similar. Intuitively, where temporal covariances reflect that alleles
associated with heritable fitness backgrounds are predictive of frequency
changes between generations, replicate covariances reflect that heritable
fitness backgrounds common to each replicate predict (under the same selection
pressures) frequency changes between replicates; we note that there is not
  a direct one-to-one correspondence between temporal and replicate
  covariances, since the latter are driven by a shared selection pressure and
  the stochastic genetic backgrounds across replicate populations. We measure
  this through a statistic similar to a correlation, which we call the
  convergent correlation: the ratio of average between-replicate covariance
  across all pairs to the average standard deviation across all pairs of
  replicates,

\begin{align}
  \label{eq:conv-corr}
  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
\end{align}
%
where $A$ and $B$ here are two replicate labels, and for the
Barghi et al. (2019) data, we use $\Delta_{_{10}} p_t$.

We have calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2}A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2}A). We find these convergent correlation coefficients
are relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while a
substantial fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result consistent with the primary finding
of the original Barghi et al. (2019) study: that alternative loci
contribute to longer term adaptation across the different replicates.

\begin{figure*}
  \centering
  \includegraphics[width=15cm]{figure-2-pnas.pdf}

  \caption{{\bf A}: The convergence correlations, averaged across
    Barghi et al. (2019) replicate pairs, through time. Each line
    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
    p_{t})$ from a starting reference generation $s$ to a later time $t$, which
    varies along the x-axis; each line corresponds to a row of the temporal
    convergence correlation matrix depicted to the right (where the
      diagonal elements represent the convergence correlations between the same
      timepoints across replicate populations). We note that convergent
      correlation for the last timepoint is an outlier; we are unsure as to the
      cause of this, e.g. it does not appear to be driven by a single pair of
      replicates. {\bf B}: The convergence correlations between individual
      pairs of replicates in the Kelly and Hughes (2019) data (note the
      confidence intervals are plotted, but are small on this y-axis scale).
      {\bf C}:  The convergence correlations between individual pairs of
      replicates in \cite{Castro2019-uk} data, for the two selection lines
      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
      two regions Castro et al. (2019) found to have signals of parallel
      selection between LS1 and LS2. Through simulations, we have found
        that the differences in convergence correlation confidence interval
        widths between these \emph{Drosophila} studies and the Longshanks study
        are due to the differing population sizes.}

  \label{fig:figure-2}
\end{figure*}

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of Kelly and Hughes (2019), which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive convergence correlations.  We find all three convergent correlation
coefficients between replicate pairs are significant (Figure
\ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$).
Additionally, we can calculate the proportion of the total variance in allele
frequency change from convergent selection pressure, analogous to our $G(t)$,
where the numerator is the convergent covariance and the denominator is the
total variance (see SI appendix, Section S4). We find that 37\% of the total
variance is due to shared allele frequency changes caused by selection (95\% CI
[29\%, 41\%]; these are similar to the convergence correlation, since the
variance is relatively constant across the replicates.

% 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859], [0.00608, 0.00834])

Next, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\cite{Marchini2014-de,Castro2019-uk}. This study includes two independent
selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
line (Ctrl) where parents were randomly selected. Consequently, this selection
experiment offers a useful control to test our convergence correlations: we
expect to see significant positive convergence correlations in the comparison
between the two Longshanks selection lines, but not between each of the control
line and Longshanks line pairs. We find that this is the case (gray confidence
intervals in Figure \ref{fig:figure-2}C), with convergence correlations between
each of the Longshanks lines to the control not being statistically different
from zero, while the convergence correlation between the two Longshanks lines
is strong (0.18) and statistically significant (CIs [0.07, 0.25]).

One finding in the Longshanks study was that two major-effect loci showed
parallel frequency shifts between the two selection lines. We were curious to
what extent our genome-wide covariances were being driven by these two outlier
large-effect loci, so we excluded them from the analysis. Since we do not know
the extent to which linkage disequilibrium around these large-effect loci
affects neighboring loci, we took the conservative precaution of excluding the
entire chromosomes these loci reside on (chromosomes 5 and 10), and
re-calculating the temporal covariances. We find excluding these large effect
loci has little impact on the confidence intervals (blue confidence intervals
in Figure \ref{fig:figure-2}C), indicating that these across-replicate
covariances are indeed driven by a large number of loci. This is consistent
with a signal of selection on a polygenic trait driving genome-wide change,
although we note that large-effect loci can contribute to the indirect change
at unlinked loci \cite{Robertson1961-ho,Santiago1995-hx}.

The presence of an unselected control line provides an alternative way to
partition the effects of linked selection and genetic drift: we can compare the
total variance in allele frequency change of the control line (which excludes
the effect of artificial selection on allele frequencies) to the total variance
in frequency change of the Longshanks selection lines. This allows us to
estimate the increase in variance in allele frequency change due to selection,
which we can further partition into the effects of selection shared between
selection lines and those unique to a selection line by estimating the shared
effect through the observed covariance between replicates (see Materials and
Methods \ref{sec:mm-partition} and SI Appendix, Section S4 for more details).
We estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
frequency change is driven by the effects of selection, of which 14\% (95\% CI
$[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\% (95\% CI
$[9\%, 23\%]$) is the effect of shared selection between the two Longshanks
selection lines.

\begin{figure}
  \centering
  \includegraphics[]{figure-3-revision.pdf}

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal
    covariances calculated in 100kb genomic windows from the
    Barghi et al. (2019) study, plotted alongside an empirical neutral null
    distribution created by recalculating the windowed covariances on 1,000
    sign permutations of allele frequency changes within tiles. The number of histogram
    bins is 88, chosen by cross validation (SI Appendix, Section
    S25). In subfigure {\bf A},
    windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
    $k=2 \times 10$ generations and in subfigure {\bf A} the covariances are
    separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal from
    the variance diagonal of the temporal covariance matrix (see cartoon of
    upper-triangle of covariance matrix in subfigures {\bf A} and {\bf B},
    where the first diagonal is the variance, and the dark gray indicates which
    off-diagonal of the covariance matrix is plotted in the histograms). {\bf
    C}: The lower and upper tail probabilities of the observed windowed
    covariances, at 20\% and 80\% quintiles of the empirical neutral null
    distribution, for varying time between allele frequency changes (i.e. which
    off-diagonal $k$). The confidence intervals are  $95\%$ block-bootstrap
  confidence intervals, and the light gray dashed line indicates the 20\% tail
probability expected under the neutral null. Similar figures for different
values of $k$ are in SI Appendix, Section S27.}

    \label{fig:figure-3} 
\end{figure}

We observed that in the longest study  we analyzed \cite{Barghi2019-qy}, some
genome-wide temporal covariances become negative at future timepoints (see the
first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
were on average going up initially are later going down in frequency, i.e.
that the average direction of selection experienced by alleles has flipped.
This might reflect either a change in the environment or the genetic
background, due to epistatic relationships among alleles altered by frequency
changes (which can occur during an optima shift; \cite{Hayward2019-kt}) or
recombination breaking up selective alleles. Such reversals in selection
dynamics could be occurring at other timepoints but the signal of a change in
the direction of selection at particular loci may be washed out when we
calculate our genome-wide average temporal covariances.  To address this
limitation, we calculated the distribution of the temporal covariances over
100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
pooling across all replicates; see SI Appendix, Fig. S26 for individuals
replicates). The covariance estimate of each genomic window will be noisy, due
to sampling and genetic drift, and the neutral distribution of the covariance
is complicated due to linkage disequilibria, which can occur over long physical
distances in E\&R and selection studies
\cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To address this, we have developed
a permutation-based procedure that constructs an empirical neutral null
distribution by randomly flipping the sign of the allele frequency changes in
each genomic window (i.e. a single random sign flip is applied to all loci in a
window). This destroys the systematic covariances created by linked selection
and creates a sampling distribution of the covariances spuriously created by
neutral genetic drift while preserving the complex dependencies between
adjacent loci created by linkage disequilibrium.  This empirical neutral null
distribution is conservative in the sense that the variances of the covariances
are wider than expected under drift alone, as selection not only creates
covariance between time intervals, but also inflates the magnitude of allele
frequency change within a time-interval.  We see (Figure \ref{fig:figure-3} A
and B) that there are an empirical excess of windows with positive covariances
between close timepoints compared to the null distribution (a heavier right
tail), and that this then shifts to an excess of windows with negative
covariances between more distant timepoints (a heavier left tail).

We quantified the degree to which the left and right tails are inflated
compared to the null distribution as a function of time, and see excesses in
both tails in Figure \ref{fig:figure-3}C. This finding is also robust to
sign-permuting allele frequency changes on a chromosome-level, the longest
extent that gametic linkage disequilibria can extend (SI Appendix, Fig. S29).
We see a striking pattern that the windowed covariances not only decay towards
zero, but in fact become negative through time, consistent with many regions in
the genome having had a reversed fitness effect at later timepoints.

\begin{figure*}
  \centering
  \includegraphics[width=15cm]{figure-4.pdf}

  \caption{Forward-in-time simulations demonstrate how temporal covariance,
    $G(t)$ trajectories, and convergence correlations arise during optima
    shifts of two different magnitudes, under Gaussian stabilizing selection.
    ({\bf A}) Trait means across 30 replicate before and after optima shifts
    (solid lines), for two different magnitudes (indicated by color). The
    optimal trait values are indicated by the purple and yellow dashed lines.
    ({\bf B}) Mean temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30
    simulation replicates, where $t$ varies along the x-axis (points), with a
    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through
    time, for 30 replicate simulations across two optima shifts. The solid line
    is a loess-smoothed average. ({\bf D}) The convergence correlations between
    two populations (each $1000$ diploids) split from a common population, that
    underwent either an optima shift in the same direction (converge) and
    opposite directions (diverge) at generation five. In subfigures ({\bf B}),
    ({\bf C}), and ({\bf D}), directional selection begins at generation five,
    when the optima shifts; this is indicated by the vertical dashed red line
    (see SI Appendix Section S8.2 for details on these
    simulations).}

  \label{fig:figure-4}
\end{figure*}

Finally we used forward-in-time simulations to explore the conditions under
which temporal and convergent correlations arise. We show a subset of our
results for a model of stabilizing selection on a phenotype where directional
selection is induced by a sudden shift in the optimum phenotype of varying
magnitudes (Figure \ref{fig:figure-4}A). We find that positive temporal
covariances are produced by such selection (Figure \ref{fig:figure-4}B), and
that these positive temporal covariances can compound together to generate a
large proportion of allele frequency change being due to selection (i.e. large
$G(t)$) over the relatively short time periods similar to our analyzed
selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of $G(t)$
increases with the strength of selection, i.e. the variance in fitness, such
that stronger selection generates larger proportions of allele frequency
change.  We find a similar picture of stronger convergent selection pressures
generating larger convergence correlations (Figure \ref{fig:figure-4}D; see
also SI Appendix, Fig. S12 for how other factors impact convergence
correlations).

Averaging across replicates, these simulation results show $G(t)$ is relatively
insensitive to the number of loci underlying the trait. However, if only a
small number of loci influence the trait, the $G(t)$ trajectories are typically
much more stochastic \emph{across} replicates. This reaffirms that the
genome-wide linked selection response we see in the Barghi et al.  (2019) data
is highly polygenic (compare Figure \ref{fig:figure-1}B to SI Appendix, Fig.
S6). Furthermore, using our simulations we find
that sampling only every 10 generations does indeed mean that our estimates of
$G(t)$ are an underestimate of the proportional effect of linked selection
as they cannot include the covariance between closely spaced generations (see
SI Appendix, Fig. S14).

Additionally, we explored other modes of selection with simulations.  We find
that the long term dynamics of the covariances under directional truncation
selection, which generates substantial epistasis, are richer than we see under
Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI Appendix,
Fig. S18). We also conducted simulations of purifying selection alone (i.e.
background selection) and find that this can also generate positive temporal
covariances (SI Appendix, Fig. S16) and under some circumstances, can even
generate convergence correlations (SI Appendix, Fig. S17).  Thus it is unlikely
that the signatures of linked selection we see are entirely the result of the
novel selection pressure the populations are exposed to, and some of this
signature may be ongoing purifying selection. Only in the case of the
Longshanks experiment, does the control line allow us to conclude that
selection that is almost entirely due to the novel selection pressure.

While none of our experiments have selected the populations in divergent
directions, in our simulations we find that such selection can generate
negative convergent correlations (Figure \ref{fig:figure-4}D). This suggests
that selection experiments combining multiple replicates, control lines, as
well as divergent selection pressures might be quite informative in
disentangling the contribution of particular selection pressures from
genome-wide allele frequency changes.

\section*{Discussion}

Since the seminal analysis of Maynard Smith and Haig (1974) demonstrating that
linked neutral diversity is reduced as an advantageous polymorphism sweeps to
fixation, over four decades of theoretical and empirical research has bettered
our understanding of linked selection.  One under-used approach to understand
the genome-wide effects of selection on polygenic trait (e.g.  on standing
variation, stems from an early quantitative genetic model of linked selection
\cite{Robertson1961-ho} and its later developments
\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
models). Implicit in these models is that autocovariance between allele
frequency change is created when there is heritable fitness variation in the
population, a signal that may be readily detected from temporal genomic data
\cite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
strong effect of linked selection may not be differentiable from genetic drift
using only single contemporary population samples or looking at temporal allele
frequency change at each locus in isolation. In this way, averaging summaries
of temporal data allows us to sidestep the key problem of detecting selection
from standing variation: that the genomic footprint leaves too soft of a
signature to differentiate from a background of genetic drift. In fact we find
that the temporal covariance signal is detectable even in the extremely
difficult to detect case of selection on highly polygenic traits
\cite{Buffalo2019-io}.

It is worth building some intuition why temporal covariance allows us to detect
such faint signals of polygenic linked selection from temporal genomic data.
Variance in allele frequency change is subject to both drift and
sampling noise, which at any single locus may swamp the temporal
covariance signal due to selection, or create spurious covariances
when selection is not acting.  However, these spurious covariances do not
share a directional signal whereas the covariances created by linked selection
do; consequently, averaging across the entire genome, the temporal signal
exceeds sampling noise.

Our analyses reveal that a sizable proportion of allele frequency change in
these experimental evolution populations is due to the (likely indirect) action
of selection. Capitalizing on replicated designs, we characterized the extent
to which convergent selection pressures lead to parallel changes in allele
frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations.  Finally, through
simulation results, we show that for a given level of additive genetic
variance, the strengths of temporal and replicate covariances depend on the
mode of selection, the details of the populations or selection experiment, and
the level of linkage disequilibria, yet the level of temporal covariance is
relatively invariant to the number of loci underlying fitness, as long as
fitness is sufficiently polygenic.

These estimates of the contribution of selection could be refined by using
patterns of linkage disequilibria (LD) and recombination which would allow us
to more fully parameterize a linked-selection model of temporal allele
frequency change \cite{Buffalo2019-io}. The basic prediction is that regions of
higher linkage disequilibrium and lower recombination should have greater
temporal autocovariance than regions with lower LD and higher recombination.
However, one limitation of these pooled sequence datasets is that none of the
studies we reanalyzed estimated linkage disequilibria data for the evolved
populations.  While there are LD data for a natural population of \emph{D.
simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a relationship
between temporal covariance and LD.  We believe this is driven by the
idiosyncratic nature of LD in evolve-and-resequence populations, which often
extends over large genomic distances \cite{Nuzhdin2013-gf,Kelly2019-dc}. Future
studies complete with LD data and recombination maps would allow one to
disentangle the influence of closely linked sites from more distant sites in
causing temporal autocovariance, and allow the fitting of more parametric
models to estimate population parameters such as the additive genetic variance
for fitness directly from temporal genomic data alone \cite{Buffalo2019-io}.
Future work could refine our $G(t)$ estimates by including selection's impact
on the variance in allele frequency terms (e.g. see equation 26 of
\cite{Buffalo2019-io}), and possibly quantifying the covariances missed when
sequencing is not done each generation; both would lead to less conservative
estimates that could show a large impact of selection.

Our primary focus here has been on evolution in laboratory populations. It is
unclear whether we should expect a similar impact of selection in natural
populations. In some of these experiments, selection pressures may have been
stronger or more sustained than in natural populations
\cite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations were
maintained at relatively small census sizes (Table \ref{table:study_summary}),
which will amplify the role of genetic drift, and increase the frequency of
rare deleterious alleles in selection lines due to founder effects. The
advantage of lab experiments is that they are closed populations; in natural
populations temporal covariance could also arise from the systematic migration
of alleles from differentiated populations. Adapting these methods to natural
populations will require either populations that are reasonably closed to
migration, or for the effect of migration to be accounted for possibly either
by knowledge of allele frequencies in source populations or the identification
of migrant individuals.

While it challenging to apply temporal methods to natural populations there is
a lot of promise for these approaches
\cite{Bergland2014-ij,Machado2018-cs}. Efforts to quantify the impact of
linked selection have found obligately sexual organisms have up to an 89\%
reduction in genome-wide diversity over long time periods
\cite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
Thus linked selection makes a sizeable contribution to long-term allele
frequency change in some species, and there is reason to be hopeful that we
could detect this from temporal data, which would help to resolve the
timescales that linked selection acts over in the wild. In our
reanalysis of the Barghi et al. (2019) study, we find evidence of complex
linked selection dynamics, with selection pressures flipping over time due to
either environmental change, the breakup of epistatic combinations or
advantageous haplotypes. Such patterns would be completely obscured in samples
from only contemporary populations. Thus, we can hope to have a much
richer picture of the impact of selection as temporal sequencing becomes more
common, allowing us to observe the effects of ecological dynamics in genomic
data \cite{Hairston2005-ga}.

Furthermore, understanding the dynamics of linked selection over short
timescales will help to unite phenotypic studies of rapid adaptation with a
detectable genomic signature, to address long-standing questions concerning
linked selection, evolutionary quantitative genetics, and the overall impact
selection has on genetic variation.

\matmethods{

\subsection*{Datasets Analyzed}

We used available genomic data data from four studies: pooled population
resequencing (pool-Seq) data from Barghi et al. (2019), Kelly and Hughes
(2019), and Bergland et al. (2014), and individual-level sequencing data from
Castro et al. (2019). In all cases, we used the variants kept after the
filtering criteria of the original studies.

\subsection*{Variance and Covariance Estimates}

To remove systematic covariances in allele frequency change caused by tracking
the reference or minor allele, we randomly choose an allele to track frequency
for each locus. Then, we calculate the variance-covariance matrix of allele
frequency changes using a Python software package we have written, available at
\url{http://github.com/vsbuffalo/cvtk}. This simultaneously calculates temporal
variances and covariances, and replicate covariances and uses the sampling
depth and number of diploid individuals to correct for bias in the variance
estimates and a bias that occurs in covariance estimates between adjacent
timepoints due to shared sampling noise (see SI Appendix, Sections S1.2, S1.3,
and S1.4 for mathematical details of these estimators). We assess that our bias
correction procedure is working adequately through a series of diagnostic plots
that ensure that the procedure removes the relationship between sampling depth
and uncorrected variance and covariances (SI Appendix, Fig. S4).  Through our
simulations we find that our estimates can differ based on how fixations and
losses are handled in long time-series (SI Appendix, Section S8.7) but none of
our findings in the main text are qualitatively altered by this decision (SI
Appendix, Figs. S19 and S20).

\subsection*{Estimating Uncertainty with a Block Bootstrap}

To infer the uncertainty of covariance, convergence correlation, and $G(t)$
estimates, we used a block bootstrap procedure. This bootstrap procedure
resamples blocks of loci, rather than individual loci, to infer the uncertainty
of a statistic in the presence of unknown correlation between loci.  As most
estimators in this paper are ratios (e.g. covariance standardize by sample
heterozygosity, $G(t)$, and the convergence correlation), which we estimate
with a ratio of averages, we exploit the linearity of expectation for efficient
computation of bootstrap samples (see SI Appendix, Fig. S3 for details).

\subsection*{Partitioning Unique and Shared Selection Effects in the Longshanks Study}
\label{sec:mm-partition}

The unselected control line in the Longshanks experiment allows us to
additionally partition the total variance in allele frequency change into
drift, shared effects of selection, and unshared effects of selection between
selected replicates. We begin by decomposing the allele frequency change in
Longshanks line 1 (LS1) as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
p_{t,\mathrm{LS}}$ where these terms are the drift in Longshanks replicate 1
($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
similarly for the Longshanks line 2, LS2). By construction, this decomposition
assumes that each of these terms are uncorrelated within replicates, so the
contribution of each term to the total variance in allele frequency change,
$\var(\Delta p_{t,\mathrm{LS1}})$, is the variance of that term's allele
frequency change.

We estimate the effects of selection by first calculating the fraction of the
total variance explained by drift. We assume the variance in allele frequency
change observed in the unselected control line ($\var(\Delta
p_{t,\mathrm{Ctrl}})$) is driven entirely by neutral genetic drift, and since
an identical breeding scheme was used across all three replicates (except
breeders for the control line were chosen at random), we can use this as an
estimate of the contribution of neutral genetic drift in the selected lines,
$\var(\Delta p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}} p_{t,\mathrm{LS1}}) =
\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$. Then, we can estimate the increase in
variance in allele frequency change due to selection as $(\var(\Delta
p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta
p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across selected lines
as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. Finally, the
covariance in allele change between replicates is used to estimate the shared
effects of selection between lines, $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
p_{t,\mathrm{LS2}}) = \var(\Delta_{_S} p_{t,\mathrm{LS}})$.

\subsection*{Windowed Covariance and the Empirical Neutral Null}

Throughout the paper, we use genomic windows for the block-bootstrap procedure.
For the \emph{D. simulans} and \emph{D.  melanogaster} data from the
Barghi et al. (2019), Kelly and Hughes (2019), and
Bergland et al. (2014) studies, we used large megabase windows for the
block bootstrap procedure, while we used a ten megabase window for the large
mouse genome data from the Castro et al. (2019) study.

Given evidence of a reversal in the direction of selection at later timepoints
in the Barghi et al. (2019) study, we calculated windowed temporal covariances
on 10 kilobase windows and looked at the distribution of these covariances
through time. We compare these distributions of windowed covariances to an
empirical neutral null created by randomly permuting the sign of allele
frequency change at the block level (to preserve the correlation structure
between loci due to LD). This destroys the systematic covariances in allele
frequency change created by linked selection, which emulates a frequency
trajectory under drift. This approach is conservative, since heritable fitness
variation also inflates the magnitude of allele frequency change more than
expected under drift, but we do not change these magnitudes. Using this
empirical neutral null distribution of windowed covariances, we calculate how
much of the observed windowed covariance distribution falls outside of
empirical null distribution for different tail probabilities. While the
comparison between the distribution of 10 kilobase windowed covariances to the
empirical neutral null created from sign-permuting 10 kilobase windows is most
natural, we wanted to ensure that our finding that the shift from mostly
positive to mostly negative windowed covariances through time (Figure
\ref{fig:figure-3}) was robust to LD extending beyond the range of these 10
kilobase windows. We took the conservative approach of also sign-permuting at
the chromosome-level, and found the same qualitative shift (SI Appendix, Fig.
S29).

\subsection*{Forward-in-time Simulations}

To explore how aspects of genetic architecture, models of selection, and
experimental design impact temporal covariance, the $G(t)$ trajectories, and
convergence correlations, we ran extensive forward-in-time simulations using
SLiM \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
simulations in Figured \ref{fig:figure-4}, but SI Appendix, Section S8
describes these simulation routines and others in detail.

We simulated directional selection on a trait by first evolving each
population of $N=1000$ diploids to equilibrium (we will refer to this as the
burnin hereafter) under GSS for $10N$ generations with the stabilizing
selection variance $V_s = 1$ and an optima set at zero. We note that the small
burnin population size means that these simulations should not be taken as
reflecting any specific natural population and they are for illustrative
purposes only.  We simulated a polygenic architecture by setting the trait
mutation rate to $10^{-8}$ per basepair, per generation, in addition to having
a separate neutral mutation of $10^{-8}$ which created neutral mutations which
we used to calculate the temporal covariances. Our simulated region was 50
megabases in length (about one quarter of a \emph{Drosophila} chromosome), and
trait alleles were randomly selected to have a $\pm 0.01$ effect size. By
tracking the trait mean through the burnin, we found it converged to the
optimum as expected.  After the burnin, the population was split into two
different replicate populations, to capture the effect of bottlenecks in
selection experiments (these population sizes varied between 50, 500, and 1000
diploids; the later representing no bottleneck). Each population then underwent
an optima shift of either 0.1, 0.5, or 1 on generation five, with the first
four generations serving as a control.  These optima shifts were either in the
same direction (converging), different directions (diverging), or only one
optima shifted (as a control). By tracking the trait mean, we saw that it
converged as expected during burnin, and the trait showed the expected
directional response to selection (SI Appendix, Fig. S7).
Using the neutral population frequency data from these simulations, we
calculated the temporal covariances, $G(t)$ trajectories, and convergence
correlations.

\subsection*{Data Availability}

All analysis was done in Python, using numpy, matplotlib, and Jupyter notebooks
\cite{Rossum1995-ni,Oliphant2006-ul,Kluyver_2016,Hunter2007-ys}; code to
reproduce these analyses is available on Github,
\url{https://github.com/vsbuffalo/cvtk/}. All data is from previous studies and
available; Barghi et al. (2019) data was downloaded from
\url{https://datadryad.org/resource/doi:10.5061/dryad.rr137kn}, Kelly and
Hughes (2019) data was downloaded from
\url{https://gsajournals.figshare.com/articles/Supplemental_Material_for_Kelly_and_Hughes_2018/7124963},
Bergland et al. (2014) data was downloaded from
\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v883p}, and Castro
et al. (2019) data was downloaded from
\url{http://ftp.tuebingen.mpg.de/fml/ag-chan/Longshanks/}.  }

\showmatmethods{}

\acknow{ We would like to thank the authors of the original studies we have
  analyzed, including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan,
  Kimberly Hughes, John Kelly, Dmitri Petrov, Campbell Rolian, Christian
  Schl{\"o}tterer. We would also like to thank Doc Edge for helpful statistical
  advice, and Dave Begun, Erin Calfee, Sarah Friedman, Andy Kern, Chuck
  Langley, Michael Turelli, Matt Osmond, Peter Ralph, and Sivan Yair for
  helpful discussions. Additionally, we thank Guy Sella and an anonymous
  reviewer whose comments greatly improved the manuscript. This research was
  supported by an NSF Graduate Research Fellowship grant awarded to VB
  (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to GC.}

\showacknow{}

\section*{Acknowledgments} 
\bibliography{biblio.bib}

\end{document}